---
output:
  bookdown::html_document2:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---


```{r echo = FALSE, cache = FALSE}
# This block needs cache=FALSE to set fig.width and fig.height, and have those
# persist across cached builds.
# source("utils.R", local = TRUE)
knitr::opts_chunk$set(fig.width=5, fig.height=4, out.width="50%")
```

Quickly Exploring Data {#CHAPTER-QUICK}
======================

Although I've used the ggplot2 package for most of the graphics in this book, it is not the only way to plot data. For very quick exploration of data, it's sometimes useful to use the plotting functions in base R. These are installed by default with R and do not require any additional packages to be installed. They're quick to type, straightforward to use in simple cases, and run very quickly.

If you want to do anything beyond very simple plots, though, it's generally better to switch to ggplot2. This is in part because ggplot2 provides a unified interface and set of options, instead of the grab bag of modifiers and special cases required in base graphics. Once you learn how ggplot2 works, you can use that knowledge for everything from scatter plots and histograms to violin plots and maps.

Each recipe in this section shows how to make a graph with base graphics. Each recipe also shows how to make a similar graph with the `ggplot()` function in ggplot2. The previous edition of this book also gave examples using the `qplot()` function from the ggplot2 package, but now it is recommended to just use `ggplot()` instead.

If you already know how to use R's base graphics, having these examples side by side will help you transition to using ggplot2 for when you want to make more sophisticated graphics.

Creating a Scatter Plot {#RECIPE-QUICK-SCATTER}
-----------------------

### Problem

You want to create a scatter plot.

### Solution

To make a scatter plot (Figure \@ref(fig:FIG-QUICK-SCATTER-BASE)), use `plot()` and pass it a vector of *x* values followed by a vector of *y* values:

```{r FIG-QUICK-SCATTER-BASE, small.mar=TRUE, fig.cap='Scatter plot with base graphics', out.width="70%"}
data("mtcars")
plot(mtcars$wt, mtcars$mpg)
```

The `mtcars$wt` returns the column named `wt` from the `mtcars` data frame, and `mtcars$mpg` is the `mpg` column.

With ggplot2, you can get a similar result using the `ggplot()` function (Figure Figure \@ref(fig:FIG-QUICK-SCATTER-GGPLOT)):

```{r FIG-QUICK-SCATTER-GGPLOT, fig.cap='Scatter plot with ggplot2', out.width="70%"}
library(ggplot2)
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point()
```

The first part, `ggplot()`, tell it to create a plot object, and the second part, `geom_point()`, tells it to add a layer of points to the plot.

The usual way to use `ggplot()` is to pass it a data frame (`mtcars`) and then tell it which columns to use for the x and y values. If you want to pass it two vectors for x and y values, you can use `data = NULL`, and then pass it the vectors. Keep in mind that ggplot2 is designed to work with data frames as the data source, not individual vectors, and that using it this way will only allow you to use a limited part of its capabilities.

